---
title: "Project"
output: html_document
---

Loading and preprocessing the Data
```{r}
library(ggplot2)

activity <- read.csv("activity.csv")

activity$date <- as.POSIXct(activity$date, tryFormats= "%Y-%m-%d")
weekday <- weekdays(activity$date)
activity <- cbind(activity,weekday)

summary(activity)
```
1. What is mean total number of steps taken per day?
```{r}
activity_total_steps <- with(activity, aggregate(steps, by = list(date), FUN = sum, na.rm = TRUE))
names(activity_total_steps) <- c("date", "steps")
hist(activity_total_steps$steps, main = "Total number of steps taken per day", xlab = "Total steps taken per day", col = "red", ylim = c(0,20), breaks = seq(0,25000, by=2500))
```

1.1 the mean and median of the total number of steps taken per day, 

```{r}
mean(activity_total_steps$steps)
median(activity_total_steps$steps)
```
2. What is the average daily activity pattern?
```{r}
average_daily_activity <- aggregate(activity$steps, by=list(activity$interval), FUN=mean, na.rm=TRUE)
names(average_daily_activity) <- c("interval", "mean")
plot(average_daily_activity$interval, average_daily_activity$mean, type = "l", col="red", lwd = 2, xlab="Interval", ylab="Average number of steps", main="Average number of steps per intervals")
```
2.2 Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
average_daily_activity[which.max(average_daily_activity$mean), ]$interval

```

3. Imputing missing values
There are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data.

we calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r}
sum(is.na(activity$steps))
```

Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

```{r}
imputed_steps <- average_daily_activity$mean[match(activity$interval, average_daily_activity$interval)]
```
Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
activity_imputed <- transform(activity, steps = ifelse(is.na(activity$steps), yes = imputed_steps, no = activity$steps))
total_steps_imputed <- aggregate(steps ~ date, activity_imputed, sum)
names(total_steps_imputed) <- c("date", "daily_steps")
```
4. Histogram of the total number of steps taken each day after missing values are imputed
```{r}
hist(total_steps_imputed$daily_steps, col = "red", xlab = "Total steps per day", ylim = c(0,30), main = "Total number of steps taken each day", breaks = seq(0,25000,by=2500))
```
4.1 Mean and median of the total number of steps taken per day:
```{r}
mean(total_steps_imputed$daily_steps)
median(total_steps_imputed$daily_steps)
```
5. Are there any differences in activity patterns between weekdays and weekends?
```{r}
is_weekday <- function(d) {
    wd <- weekdays(d)
    ifelse (wd == "Saturday" | wd == "Sunday", "weekend", "weekday")
}

wx <- sapply(activity$date, is_weekday)
activity$datetype <- as.factor(wx)
head(activity$date)
```
```{r}
activity_by_date <- aggregate(steps ~interval+datetype, data=activity, FUN=mean)

library(lattice)
xyplot(steps ~ interval | factor(activity$datetype),
       layout = c(2, 1),
       xlab="Interval",
       ylab="Number of steps",
       type="l",
       lty=1,
       data=activity_by_date)
```

